---
title: 'Lab 4- Jayant Shah'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Data:

```{r}
library(readxl)
# Read Defensive
defense = read_excel('C:/Users/Jayant Shah/Downloads/2020 data file.xlsx', sheet = 'Defensive')
str(defense)
# Read Offensive
offense = read_excel('C:/Users/Jayant Shah/Downloads/2020 data file.xlsx', sheet = 'Offensive')
str(offense)
# Read Biography
biography = read_excel('C:/Users/Jayant Shah/Downloads/2020 data file.xlsx', sheet = 'Biography')
str(biography)
```

# Part One

1. 

```{r}
defense$Name = as.factor(defense$Name)
defense$Opponent_Opponent = as.factor(defense$Opponent_Opponent)
offense$Name = as.factor(offense$Name)
offense$Opponent_Opponent = as.factor(offense$Opponent_Opponent)
```

2.

```{r}
library(dplyr)
defClean = defense %>% mutate(across(where(is.character), as.numeric))
offClean = offense %>% mutate(across(where(is.character), as.numeric))
biography$Weight = as.numeric(biography$Weight)
```

3. 

```{r}
# Height is measured in foot
biography$Height = as.numeric(gsub("-", ".", biography$Height))
bioClean = biography
```

4.

```{r}
defClean$game = ifelse(defClean$Opponent_Opponent == "Oklahoma St.", "second", "first")
offClean$game = ifelse(offClean$Opponent_Opponent == "Oklahoma St.", "second", "first")
```

The structures of the data frames

defClean:

```{r}
str(defClean)
```

offClean:

```{r}
str(offClean)
```


bioClean:

```{r}
str(bioClean)
```

# Part Two

1.

```{r}
library(tidyr)
offClean_tydi = offClean[,1:12] %>%
  pivot_longer(!c(Name, Opponent_Opponent), names_to = "stat", values_to = "count")
head(offClean_tydi,10)
```

2. 

```{r}
df2_2 = offClean_tydi %>% group_by(Name, stat) %>% summarise(n=sum(count))
head(df2_2,10)
```

3.

Passing_CMP-ATT distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Passing_CMP-ATT")$n, na.rm = T)>0
```

Here any observations not equal to NA.

Passing_INT distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Passing_INT")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Passing_INT")$n, main = "Distribution of Passing_INT", xlab = "Passing_INT", breaks = 10)
```

Passing_TD distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Passing_TD")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Passing_TD")$n, main = "Distribution of Passing_TD", xlab = "Passing_TD", breaks = 10)
```

Passing_YDS distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Passing_YDS")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Passing_YDS")$n, main = "Distribution of Passing_YDS", xlab = "Passing_YDS", breaks = 10)
```

Receiving_REC distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Receiving_REC")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Receiving_REC")$n, main = "Distribution of Receiving_REC", xlab = "Receiving_REC", breaks = 10)
```

Receiving_TD distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Receiving_TD")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Receiving_TD")$n, main = "Distribution of Receiving_TD", xlab = "Receiving_TD", breaks = 10)
```


Receiving_YDS distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Receiving_YDS")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Receiving_YDS")$n, main = "Distribution of Receiving_YDS", xlab = "Receiving_YDS", breaks = 10)
```

Rushing_ATT distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Rushing_ATT")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Rushing_ATT")$n, main = "Distribution of Rushing_ATT", xlab = "Rushing_ATT", breaks = 10)
```

Rushing_TD distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Rushing_TD")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Rushing_TD")$n, main = "Distribution of Rushing_TD", xlab = "Rushing_TD", breaks = 10)
```

Rushing_YDS distribution:

```{r}
sum(subset(df2_2, df2_2$stat=="Rushing_YDS")$n, na.rm = T)>0
```
Here are observations not equal to NA.

```{r}
hist(subset(df2_2, df2_2$stat=="Rushing_YDS")$n, main = "Distribution of Rushing_YDS", xlab = "Rushing_YDS", breaks = 10)
```


There are no Passing_CMP-ATT observations. Only one non-zero value each for Passing_INT, Passing_TD and Passing_YDS.

4. 

```{r}
df2_4_Oregon = subset(offClean, offClean$Opponent_Opponent == "Oregon")
df2_4_Oklahoma = subset(offClean, offClean$Opponent_Opponent == "Oklahoma")
plot(df2_4_Oregon$Receiving_YDS, ylab = "Receiving_YDS", main = "Opponent Oregon")
plot(df2_4_Oklahoma$Receiving_YDS, ylab = "Receiving_YDS", main = "Opponent Oklahoma")
```

ISU have better offense against Oklahoma.

## Now turn attention to the player biography.

4.

```{r}
bioClean_sep = bioClean %>% separate(Hometown, c("City","State"), sep = "([,])")
head(bioClean_sep, 10)
```

5.

```{r}
bioClean_sep %>% group_by(State) %>% summarise(number_of_players=n())
```

6.

The performance of quarterback Brock Purdy:
 
```{r}
subset(offClean_tydi, offClean_tydi$Name == "Purdy, Brock") %>% group_by(stat) %>% summarise(n=sum(count))
```


